CannonRV: Measuring radial velocities with a data-driven spectral model

### 2016-12-04 paper/ms.tex
Compare two methods
Compare two formula for radial velocity shifts

$\sigma_{jn}^2$ is the data uncertainty for star n at pixel j. The number of stellar and pixel each star are N and J. Here N and J are the number of the stellar and the number of pixel each star.  Write the objective function in the form of matrix can make it simpler. There will be N objective functions and N sets of a, b and c. So we introduce several matrix: $Y_n$, $A_n$, $C_n$ and $X_n$, which are defined as followed:

@@

\begin{equation}
X_n = [\sum_{n=1}^{n=N}A_n^TC_n^{-1}A]^{-1} \cdot [\sum_{n=1}^{n=N}A^TC_n^{-1}Y_n]
\end{equation}

@@

We will do this optimization for each visit of every star in APOGEE DR10. The spectrum we use is in apstar file. We will re-analyze the velocities after the optimization and deliver back the velocities correction for the APOGEE data.
\bigskip

@@

To begin with, we set the radial velocity to be positive if the star is leaving us. If a=1, b=0 and c=0, the optimized spectrum moves one pixel left from the inferred spectrum, which means the data spectrum should move one pixel right and the radial velocity of the data spectrum is not zero and it has a blue shift. Thus the radial velocity correction is negative and can be obtained by using (8). It's more convenient to use the logarithmic wavelength.
\begin{equation}
\log_{10}\lambda_0 - \log_{10}\lambda_s = \log_{10}(1+\beta)
\end{equation}

@@

Thus we can use a, b and c to determine the radial velocity correction for the combined spectrum and each visit of the APOGEE data.
\begin{equation}
v_{deviation} = (c-a)*v_{pixel}
\end{equation}
Where $v_{deviation}$ is the deviation of the APOGEE data and $v_{pixel}$ =  4144.68 m/s.

@@

@@

Here we choose the four stars with biggest a-c and obtain the radial velocity corrections for them. The spectrum we use are continuum-normalized combined spectrum of these four stars. Then we derive their radial velocity corrections and de-Doppler the wavelengths. Use Sinc interpolation to resample them onto the same wavelength scale and obtain the calibrated spectrum. The wavelength scale is the same as the one in Apstar file, which is described in the former section. Finally, plot the data spectrum and the calibrated spectrum in figure 11.

@@

\centerline{Fig.11 Comparison of the calibrated spectrum and the data spectrum for the four stars with biggest a-c}

\centerline{ The black line is the data spectrum. The green line is the inferred spectrum and the red line is the calibrated data spectrum}
\bigskip

We do the same thing four the four stars with biggest c-a.

@@

\centerline{Fig.12 Comparison of the calibrated spectrum and the data spectrum for the four stars with biggest c-a.}

\centerline{ The black line is the data spectrum. The green line is the inferred spectrum and the red line is the calibrated data spectrum}
\bigskip

After that, we want to calibrate the spectrum for each visit. To begin with, we need to check whether our normalization works in each visit. The method is to infer labels for each visit and compare them with the ones from the combined spectrum. If our normalization works, the inferred labels from each visit for the same star should be similar to the ones from the combined spectrum. We plot the histogram which shows the deviations of the inferred labels of each visit from the ones generated by the combined spectrum of the same star. The data we use is the 614 stars in experiment 1 to 4. If most of the deviations are small, the normalized spectrum for each visit doesn't lose shape, which means our normalization method works.

@@

\centerline{Fig.13 Histogram for Delta Teff}
\bigskip

@@

\centerline{Fig.14 Histogram for Delta logg}

@@

\centerline{Fig.15 Histogram for Delta Fe/H}
\bigskip

@@

\centerline{Fig.16 Comparison of each visit for the first star in figure 8.}
\bigskip

@@

>>>>

$\sigma_{jn}^2$ is the data uncertainty for star n at pixel j. The number of stellar and pixel each star are N and J. Here N and J are the number of the stellar and the number of pixel each star.  Write the objective function in the form of matrix can make it simpler. There will be N objective functions and N sets of a, b and c. So we introduce several matrix: $Y_n$, $A_n$, $C_n$ and $X_n$ for star n, which are defined as followed:

@@

\begin{equation}
X_n = [A_n^TC_n^{-1}A_n]^{-1} \cdot [A_n^TC_n^{-1}Y_n]
\end{equation}

@@

We will do this optimization for each visit of every star in APOGEE DR10. The spectrum we use is in apstar file. We will re-analyze the velocities after the optimization and deliver back the velocities shift for the APOGEE data.
\bigskip

@@

To begin with, we set the radial velocity to be positive if the star is leaving us. If a=1, b=0 and c=0, the optimized spectrum moves one pixel left from the inferred spectrum, which means the data spectrum should move one pixel right and the radial velocity of the data spectrum is not zero and it has a blue shift. Thus the radial velocity shift is negative and can be obtained by using (8). It's more convenient to use the logarithmic wavelength.
\begin{equation}
\log_{10}\lambda_0 - \log_{10}\lambda_s = \log_{10}(1+\beta)
\end{equation}

@@

Thus we can use a, b and c to determine the radial velocity shift for the combined spectrum and each visit of the APOGEE data.
\begin{equation}
v_{shift} = {{(c-a)*v_{pixel}}\over{4\cdot(a+c-2b)}}
\end{equation}
Where $v_{shift}$ is the shift of the radial velocity in APOGEE data and $v_{pixel}$ =  4144.68 m/s.
\bigskip

@@

\centerline{ The black line is the observed spectrum. The green line is the inferred spectrum and the red line is the optimized spectrum.}

@@

Here we plot the histogram of the radial velocity shifts in figure 11 and the radial velocity shifts in figure 12. The data we use is the combined spectrums of the 614 stars in the former experiment.

@@

\centerline{Fig.11 Histogram of the radial velocity shifts in combined spectrums for the 614 stars. }
\centerline{We exclude the outliers which have velocity shifts bigger than 2000 m/s}
\centerline{The formula to derive velocity shifts is $v_{shift} = {{(c-a)*v_{pixel}}\over{4\cdot(a+c-2b)}}$}
\bigskip

@@

\centerline{Fig.12 Radial velocity shifts for combined spectrums of the 614 stars against mean fiber number}
\centerline{We exclude the outliers which have velocity shifts bigger than 2000 m/s}
\centerline{The formula to derive velocity shifts is $v_{shift} = {{(c-a)*v_{pixel}}\over{4\cdot(a+c-2b)}}$}
\bigskip

We also plot the four stars with biggest radial velocity shifts, which are in figure 13.

@@

\centerline{Fig.13 Spectrum of the four stars with biggest radial velocity shifts.}

\centerline{ The black line is the observed spectrum. The green line is the inferred spectrum and the red line is the optimized spectrum.}

After that, we want to calibrate the radial velocity for each visit. To begin with, we need to check whether our normalization works in each visit. The method is to infer labels for each visit and compare them with the ones from the combined spectrum. If our normalization works, the inferred labels from each visit for the same star should be similar to the ones from the combined spectrum. We plot the histogram which shows the deviations of the inferred labels of each visit from the ones generated by the combined spectrum of the same star. The data we use is the 614 stars in experiment 1 to 4. If most of the deviations are small, the normalized spectrum for each visit doesn't lose shape, which means our normalization method works.

@@

\centerline{Fig.14 Histogram for Delta Teff}
\bigskip

\centerline{Fig.15 Histogram for Delta logg}

@@

\centerline{Fig.16 Histogram for Delta Fe/H}
\bigskip

@@

\includegraphics[width=170mm]{figure_17.png}
\bigskip

\centerline{Fig.17 Comparison of each visit for the first star in figure 8.}
\bigskip

@@

I plot the histogram of radial velocity shifts derived by using $v_{shift} = (c-a)*v_{pixel}$ in diagnostic figure 2 and velocity shifts against mean fiber number in diagnostic figure 3.

\includegraphics[width=170mm]{d_figure_2.png}
\bigskip

\centerline{Diagnostic Fig.2 Histogram of radial velocity shifts}
\centerline{The formula to derive velocity shifts is $v_{shift} = (c-a)*v_{pixel}$}
\bigskip

\includegraphics[width=170mm]{d_figure_3.png}
\bigskip

\centerline{Diagnostic Fig.3 Radial velocity shifts against mean fiber number}
\centerline{The formula to derive velocity shifts is $v_{shift} = (c-a)*v_{pixel}$}
\bigskip

We also plot the four stars with biggest radial velocity shifts, which are in figure 13. The formula I use is $v_{shift} = (c-a)*v_{pixel}$.

\includegraphics[width=170mm]{d_figure_4.png}
\bigskip

\centerline{Diagnostic Fig.4 Spectrum of the four stars with biggest radial velocity shifts.}

\centerline{ The black line is the observed spectrum. The green line is the inferred spectrum and the red line is the optimized spectrum.}

The original histogram of radial velocity shifts for figure 11.
\bigskip

\includegraphics[width=170mm]{d_figure_5.png}
\bigskip

\centerline{Diagnostic Fig.5 Histogram of radial velocity shifts}
\centerline{The formula to derive velocity shifts is $v_{shift} = {{(c-a)*v_{pixel}}\over{4\cdot(a+c-2b)}}$}
\bigskip

\includegraphics[width=170mm]{d_figure_6.png}
\bigskip

\centerline{Diagnostic Fig.6 Radial velocity shifts against mean fiber number}
\centerline{The formula to derive velocity shifts is $v_{shift} = {{(c-a)*v_{pixel}}\over{4\cdot(a+c-2b)}}$}
\bigskip
